---
title: "Week5 code&note"
author: "Franklin"
date: "2022/3/21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Required packages
```{r}
library(tidyverse)
```
## Basic Concepts of Probability
* **Probability** is the likelihood that an outcome occurs.  
Probabilities are expressed values between 0 and 1.
* An **experiment** is the process that results in an outcome. In a random experiment, outcome cannot be predicted with certainty
* The **outcome** of an experiment is a result that we observe
* The **sample space** is the collection of all possible outcomes of an experiment
* An **event** is a collection of one or more outcomes from a sample space

## Probability Definitions
Probabilities may be defined from one of three perspectives:    
1. *classical definition*: Probabilities can be deduced(推导) from theoretical(理论的) arguments.
2. *Relatice frequency definition*: Probabilities are based on empirical(经验的) data
. *Subjective definition*: Probabilities are based on judgement and experience

## Basic Concepts of Probability
* Probability of any event is the sum of probability of outcome that comprise that event
* Probability of complement of an event A is $P(A_{complent}) + P(A) = 1$ 
* If A and B are mutually exclusive, then $P(A \space or\space B) = P(A) + P(B)$
* If two events are not mutually exclusive, then $P(A\space or\space B) = P(A) + P(B) - P(A \space and \space B)$

## Conditional Probability
* $P(A|B) = p(A&B)/P(B)$, the condional probability of A given B

## Random Variables
A random variable is a numerical description of the outcome of an experiment  
* A discrete random variable is one for which teh number of possible outcomes can be counted
* A continuous random variable has outcomes inf, over one or more continuous intervals or real numbers

## Probability distribution
A characterization of the possible values that a **Random Variable** may assume along with the probability of assuming these values  
**Representation:**  
* X axis = all possible values
* Y axis = probability of observing each value
* The sum of area under the function is always 1

## Probability Distribution
### 1. Discrete Probability Distribution
Called Probability Mass Function(PMF), has the following characteristic:  
1. $f_{X}(x), x_{i} \space represents \space the \space i^{th} \space value \space of \space X, \space 0 < f(x_{i}) < 1$ for all i.  
2. $\sum_{i} f(x_{i}) = 1$  
3. For cumulative distribution function(CDF), $F(x) = P(X<=x)$  
4. Some Formula of CDF:  
  * Pr(X > x) = 1 - F(x)， 因为定义，F(x) = P(X<=x), {x:X<=x} 和 {x: X>x} 是互补的事件, 所以必然有 Pr(X > x) = 1 - F(x)  
  * $Pr(x_{1}<X<=x_{2}) = F(x_2) - F(x_1)$同上，通过三个集合来证明即可: A = {X: X<=x1}, B = {X: x1 < X <=x2}, C = {X: X <=x2}则有:    
  $Pr(B)\space=Pr(C) \space - Pr(A \cap B) = Pr(C) \space - \space Pr(A)$

### 2. Expected Value of a Discrete Random variable（期望）
Has formula:  
$ E[X] = \sum x_{i}f(x_i)$, 即高中的分布,其中, xi 为实际的值，f(xi)概率值(CDF)  
### 3. Variance of a discrete random variable
Has formula:  
$Var[X] = \sum (x_{i} - E[X])^2f(x_{i})$, 即用期望估计均值，同时乘上对应的概率

### 3. Bernoulli Distribution 伯努利分布
其PMFformula:  
f(x) = p, if x = 1, else 0, if x=0  
E[X] = p, Var[X] = p(1-p)  

### 4. Binomial Distribution 二项分布
不赘述了，见ppt
### 5. Poisson Distribution
同上，见ppt

## Continuous Probability Distributions
**Probability density function (PDF)**  
A curve described by a mathematical function that characterizes a continuous random variable  
Properties of PDF:  
* f(x) >=0 for all values of x
* Total area under the pdf is 1
* P(X=x) = 0, 不能有一个确定的值，其实际公式为: cdf: Fx(x) - Fx左极限(x)(对于continuous, 其一定为0)
* probabilities are only defined over an interval
* P(a<=x<=b) = F(b) - F(a)

### uniform distribution
见ppt
### Discrete Uniform Distribution
A variation of the uniform distribution is one for which the random variable is restricted to integer values between a and b, this is called a discrete uniform distribution.

### Normal Distribution
**Properties**
1. symmetric
2. Mean = median = mode
3. Range is unbounded
4. Empirical rules apply($+,- \space 3\sigma$)  

**Using Standard normal distribution Tables**
Change any non-standard normal distribution to standard:  
By using z-score:  
$z = (x-u)/\sigma$

### Exponential Distribution
* Models the time between randomly occurring events
* Density function:  
$f(x)\space=\space\lambda e^{-\lambda x}$,for x>=0  
* CDF:  
$F(x)=1-e^{-\lambda x}$  
If the number of events occurring during an interval of time has a Possion distribtion, then the time between events is exponentially distributed.

```{r}
# representation of different norm
# dnorm, represent as normal distribution pdf
dnorm(0, mean = 0, sd = 1)
# pnorm, p 表示probability，其返回的是积分
pnorm(0)# calculate the cumulative probability
# qnorm, the CDF inverse function, 这里的q指的是quantile，即分位数
# 即计算的是标准正态分布中, 第50百分位数的Z-score是多少
qnorm(0.5)
# rnorm， 生成一组符合正态分布的随机数
rnorm(5)
# An example
x <- seq(from = -3, to =3, by = 0.1)
y = dnorm(x)
data1 <- data.frame(x=x,y=y)
ggplot(data1, aes(x,y)) + 
  geom_line() + 
  labs(x = "Normal Deviate", y="Density") +
  scale_x_continuous(breaks = seq(-3,3,1))
```
## Statistical Sampling
1. Sampling is the foundation of statistical analysis
2. Estimator are measures used to estimate unknown population parameters
3. A point estimate is a single number derived from a sample that is used to estimate the value of a population parameters

## Sampling Error
1. Sampling error occurs because samples are only a subset of the total population.(Sampling error depends on the size of the sample relative to the population. This type of sampling error cannot be totally avoided)  
2. Non-sampling error occurs when the sample does not adequately represent the target population.(Nonsampling error usually results from a poor sample design or choosing the wrong population frame)

## Sampling distributions
* Sampling distribution of the mean is the distribution of the means of all possible samples of a fixed size n from some population
* The standard deviation of the sampling distribution of the mean is called the standard error of the mean,即为样本抽样误差的指标
* As n increase, the standard error is decreases

## Central limit theorem
1. 如果samplesize足够大， 则sampling distribution 的均值是接近于服从正态分布的，其有一个均值等于总体均值
2. 如果population是正态分布的，则样本分布对于任何的采样size，都是正态分布的.
3. Central limit theorem allow us to use the theory we learned about calculating probabilities for normal distributions to draw conclusions about sample means.

## Example
The purchase order order amounts for books on a publisher’s website is  normally distributed with a mean of $36 and a standard deviation of $8.

Find the probability that:  
1. someone’s purchase amount exceeds $40.
2. the mean purchase amount for 16 customers exceeds $40
```{r}
#1. for question 1, using pnorm to find the probability 
y = 1 - pnorm(q = 40, mean = 36,sd = 8)
y
#2. it's the mean distribution, so we have u_xbar = u_population
# std_xbar = std_population/sqrt(n)
y2 = 1 - pnorm(q =40, mean = 36, sd = 2)
y2
```
## Interval Estimates
1. An interval estimate provides a range for a population characteristic based on a sample
2. 100(1-a)% probability interval is any interval [A,B] s.t. the probability of falling between A and B is 1-a
3. Notice that Probability intervals are often centered on the mean or median

## Confidence interval
见回归笔记
